---
title: "ROC Curves"
output: html_document
---

Use `ROCR` package to produce ROC Curves with a custom short function and load all `svm` packages

```{r, results='hide'}
library(ROCR)
library(e1071)
```

```{r}
rocplot=function(pred, truth, ...) {
    predob = prediction(pred, truth)
    perf = performance(predob, "tpr", "fpr")
    plot(perf,...)
}
```

Simulate the data and the data frame

```{r}
set.seed(1)
x=matrix(rnorm(200*2), ncol=2)

# To seperate X values in non-linear way
x[1:100,] <- x[1:100,] + 2
x[101:150,] <- x[101:150,] - 2

y <- c(rep(1,150), rep(2,50))
dat <- data.frame(x=x, y=as.factor(y)) # only works with a data frame
train <- sample(200, 100) # 100 samples between 0 to 200, good way to get the training
```

In order to obtain the fitted values for a given SVM model fit, we use decision.values=TRUE when fitting svm(). Then the `predict()` function will output the fitted values. This can apply to both SVMs and support vector classifiers.

You just create a training dataset.

```{r}
svmfit.opt <- svm(y~., data=dat[train,], 
                  kernel="radial",gamma=2, cost=1,decision.values=T)
fitted <- attributes(predict(svmfit.opt,dat[train,],
                             decision.values=TRUE))$decision.values
```
Now we can produce the ROC plot.

```{r}
# par(mfrow=c(1,2))
rocplot(fitted,dat[train,"y"], main="Training Data")
```

SVM appears to be producing accurate predictions. By increasing $\gamma$, we can produce a more flexible fit and generate further improvements in accuracy.

You need to include the previous code in this block as well.

```{r}
rocplot(fitted,dat[train,"y"], main="Training Data")

svmfit.flex <- svm(y~., data=dat[train,], 
                   kernel="radial",gamma=50, cost=1, decision.values=T)
fitted <- attributes(predict(svmfit.flex,dat[train,],decision.values=T))$decision.values
rocplot(fitted, dat[train,"y"], add=T, col="red")
```

The red is basically perfect but maybe overfit as it is only on the training data. We perfer prediction accuracy on the test data.

use model with $\gamma=2$ appears to provide the most accurate results.


```{r}
fitted <- attributes(predict(svmfit.opt,dat[-train,],
                             decision.values=T))$decision.values
rocplot(fitted,dat[-train,"y"],main="Test Data")

fitted <- attributes(predict(svmfit.flex,dat[-train,],
                             decision.values=T))$decision.values
rocplot(fitted,dat[-train,"y"],add=T,col="red")
```

I think the ROC curve is just to tell us if the model is good enough for the testing data by playing with `gamma`.