---
title: "Baseball salary prediction model using Best Subset Selection"
output: html_document
---

## Synopsis

The idea is to apply the best subset selection approach and create a model that predicts a baseball player's Salary on the basis of various statistics associated with performance in the previous year.

## Analysis

Since the `Salary` variable is missing from some of the players. `is.na()` function is used to identify the missing observations.

### Missing value

```{r}
library(ISLR)

 # fix(Hitters)
names(Hitters) # variablenames 

dim(Hitters)
sum(is.na(Hitters$Salary)) # total number of players without a recorded salary
```

Hence we see that Salary is missing for 59 players. The `na.omit()` function removes all of the rows that have missing values in any variable.

```{r}
Hitters=na.omit(Hitters)
dim(Hitters)
sum(is.na(Hitters))
```

We are omitting all the NA

### Subset Selection 

Use `regsubsets()` to perform best subregsubsets() set selection by identifying the best model that contains a given number of predictors.


```{r}
library(leaps)
regfit.full <- regsubsets(Salary~.,Hitters)
summary(regfit.full)
```

The asterisk indicates that a given variable is included in the corresponding model.

This output indicates that the best two-variable model contains only `Hits` and `CRBI`. 


### Subset Selection Full Model

By default, regsubsets() only reports results up to the best eight-variable model. But the nvmax option can be used in order to return as many variables as are desired. Here we fit up to a 19-variable model.


### Examine $R^2$, $RSS$, adjusted $R^2$, $C_p$, and $BIC$ on plots


```{r}
regfit.full <- regsubsets(Salary~.,data=Hitters,nvmax=19)
reg.summary <- summary(regfit.full)
```

The `summary()` function also returns $R^2$, $RSS$, adjusted $R^2$, $C_p$, and $BIC$.

We examine these to try to select the best overall model. 

```{r}
names(reg.summary)
```

For instance, we see that the $R^2$ statistic increases from 32%, when only one variable is included in the model, to almost 55%, when all variables are included. As expected, the $R^2$ statistic increases monotonically as more variables are included.

So, more variables included, the more accurate the model it is. The plots indicate the same was well, where $RSS$ and Adjusted $R^2$ 

```{r}
reg.summary$rsq
```

```{r}
par(mfrow=c(1,2))
plot(reg.summary$rss, xlab="Number of Variables", ylab="RSS", type="l")
plot(reg.summary$adjr2, xlab="Number of Variables", ylab="Adjusted RSq", type="l")
```

The `which.max()` function can be used to identify the location of the maximum point of a vector. We will now plot a red dot to indicate the model with the largest adjusted $R^2$ statistic.

```{r}
par(mfrow=c(1,2))
plot(reg.summary$rss, xlab="Number of Variables", ylab="RSS", type="l")
plot(reg.summary$adjr2, xlab="Number of Variables", ylab="Adjusted RSq", type="l")
which.max(reg.summary$adjr2)
points(11, reg.summary$adjr2[11], col="red", cex=2, pch=20)
```


In a similar fashion we can plot the $C_p$ and $BIC$ statistics, and indicate the models with the smallest statistic using `which.min()`

```{r}
par(mfrow=c(1,2))
plot(reg.summary$cp, xlab="Number of Variables", ylab="Cp", type='l')
which.min(reg.summary$cp)
points(10,reg.summary$cp[10],col="red",cex=2,pch=20)
which.min(reg.summary$bic)
plot(reg.summary$bic,xlab= "Number of Variables",ylab="BIC", type='l')
points(6,reg.summary$bic[6],col="red",cex=2,pch=20)
```

### Choose best model from full Subset selection model

Use `regsubsets()` function of its built-in `plot()` command to display the selected variables for the best model with a given number of predictors, ranked according to the $BIC$, $C_p$, adjusted $R^2$, or $AIC$. 

```{r}
plot(regfit.full,scale="r2")
plot(regfit.full,scale="adjr2")
plot(regfit.full,scale="Cp")
plot(regfit.full,scale="bic")
```

The top row of each plot contains a black square for each variable selected according to the optimal model associated with that statistic.

For instance, we see that several models share a $BIC$ close to -150

However, the model with the lowest $BIC$ is the six-variable model that contains only `AtBat`, `Hits`, `Walks`, `CRBI`, `DivisionW`, and `PutOuts`. This is the top row.

We can use the `coef()` function to see the coefficient estimates associated with this model.


```{r}
coef(regfit.full,6)
```

So even we are using the full model with 19 variables, Best Selection technique has been able to choose the best model with 6 variables.