---
title: "Baseball Salary Prediction using Forward and Backward Stepwise Selection"
output: html_document
---

## Synopsis

The idea is to apply the best Forward and Backward Stepwise selection approach and create a predictive model model on baseball player's Salary on the basis of various statistics associated with performance in the previous year.


## Analysis

### Missing value

```{r}
library(ISLR)
# names(Hitters) # variablenames 
# dim(Hitters)
# sum(is.na(Hitters$Salary)) # total number of players without a recorded salary
Hitters=na.omit(Hitters)
# dim(Hitters)
# sum(is.na(Hitters))
```

### Forward Backward Selection 

Use `regsubsets()` function to perform forward stepwise or backward stepwise selection, using the argument `method="forward"` or `method="backward"`.

```{r}
library(leaps)
regfit.fwd <- regsubsets(Salary~., data=Hitters, nvmax=19, method="forward")
summary(regfit.fwd)
regfit.bwd <- regsubsets(Salary~., data=Hitters, nvmax=19, method="backward")
summary(regfit.bwd)
```

We see that using forward stepwise selection, the best one variable model contains only `CRBI`, and the best two-variable model additionally includes `Hits`. `CRBI` has asterisk for all 19 variables where `Hits` makes up 18 variables. Ultimately, we should really depend on coefficients for model selection.

For this data, the best one-variable through six variable models are each identical for best subset and forward selection. So, both models work just in the same way.

However, the best seven-variable models identified by forward stepwise selection, backward stepwise selection, and best subset selection are different.


### Coefficient for Best model selection

We have used 3 techniques to build our models but they don't always agree as indicated in the 7-variable models.

```{r}
regfit.full <- regsubsets(Salary~.,data=Hitters,nvmax=19)
```


```{r}
coef(regfit.full,7) # Full Best model selection
coef(regfit.fwd,7) # forward stepwise
coef(regfit.bwd,7) # backward stepwise
```

So ultimately, It is OK to which one we choose to use.
