---
title: "Predict Salary Using Lasso"
output: html_document
---

## Synopsis

We will predict baseball players' salary using Lasso.

We saw that ridge regression with a wise choice of $\lambda$ can outperform least squares as well as the null model on the Hitters data set. We now ask whether the lasso can yield either a more accurate or a more interpretable model than ridge regression.

In order to fit a lasso model, we once again use the glmnet() function with the use of argument `alpha=1`.

## Analysis

### Handle missing data

```{r}
library(ISLR)
Hitters=na.omit(Hitters)
```

### Manage dataset in matrix for Lasso

```{r}
x=model.matrix(Salary~.,Hitters)[,-1]
y=Hitters$Salary
```

### Generate training and testing data for Lasso

```{r}
set.seed(1)
train=sample(1:nrow(x), nrow(x)/2)
test=(-train)
y.test=y[test]
```


### Build Lasso Model

```{r, message=FALSE, warning=FALSE}
library(glmnet)
grid <- 10^seq(10, -2, length=100) # alpha=0 
lasso.mod=glmnet(x[train,],y[train],alpha=1,lambda=grid)
plot(lasso.mod)
```

We can see from the coefficient plot that depending on the choice of tuning parameter, some of the coefficients will be exactly equal to zero.

### Perform Cross-Validation

Perform cross-validation and compute the associated test error.

```{r}
set.seed(1)
cv.out=cv.glmnet(x[train,],y[train],alpha=1)
plot(cv.out)
bestlam=cv.out$lambda.min
lasso.pred=predict(lasso.mod,s=bestlam,newx=x[test,])
mean((lasso.pred-y.test)^2)
```

This is substantially lower than the test set MSE of the null model and of least squares, and very similar to the test MSE of ridge regression with $\lambda$ chosen by cross-validation.

However, the lasso has a substantial advantage over ridge regression in that the resulting coefficient estimates are sparse. Here we see that 12 of the 19 coefficient estimates are exactly zero. So the lasso model with $\lambda$ chosen by cross-validation contains only seven variables.

### Choose Cross-Validation 

```{r}
out=glmnet(x,y,alpha=1,lambda=grid)
lasso.coef=predict(out,type="coefficients",s=bestlam)[1:20,]
lasso.coef
lasso.coef[lasso.coef!=0]
```


