---
title: "Auto fitting with Leave-One-Out Cross-Validation"
output: html_document
---

## Synopsis

The LOOCV estimate can be automatically computed for any generalized linear model.


## Analysis

```{r}
library(ISLR)
set.seed(1)
train=sample(392,196)
```

we used the `glm()` function to perform logistic regression by passing in the `family="binomial"` argument. 

But if we use `glm()` to fit a model without passing in the family argument, then it performs linear regression, just like the `lm()` function.

### GLM function

```{r, cache=TRUE}
glm.fit=glm(mpg~horsepower,data=Auto)
coef(glm.fit)
```

### LM function

```{r, cache=TRUE}
lm.fit=lm(mpg~horsepower,data=Auto)
coef(lm.fit)
```

Both yield identical linear regression models.

We will perform linear regression using the `glm()` function rather than the `lm()` function because the latter can be used together with `cv.glm()`. The `cv.glm()` function is part of the `boot` library.

```{r, cache=TRUE}
library(boot)
glm.fit=glm(mpg~horsepower,data=Auto)
cv.err=cv.glm(Auto,glm.fit)
cv.err$delta
```

The `cv.glm()` function produces a list with several components. The two numbers in the delta vector contain the cross-validation results.

In this case case the numbers are identical (up to two decimal places) and correspond to the LOOCV statistic. 

Our cross-validation estimate for the test error is approximately 24.23.

### Polynomial Fit

We can repeat this procedure for increasingly complex polynomial fits.

To automate the process, we use the for() function to initiate a for loop which iteratively fits polynomial regressions for polynomials of order i = 1 to i = 5, computes the associated cross-validation error, and stores it in the ith element of the vector cv.error. We begin by initializing the vector. This command will likely take a couple of minutes to run.


```{r, cache=TRUE}
library(boot)
cv.error=rep(0,5)

for (i in 1:5){
 glm.fit=glm(mpg~poly(horsepower,i),data=Auto)
 cv.error[i]=cv.glm(Auto,glm.fit)$delta[1]
}

cv.error
```

We see a sharp drop in the estimated test MSE between the linear and quadratic fits, but then no clear improvement from using higher-order polynomials.

