---
title: "Auto Fitting with Bootstrap"
output: html_document
---

## Synopsis

To use Boostrap to analyse Auto dataset and estimate the accuracy of the linear regression model on the dataset.


## Analysis

### Estimating the Accuracy of a Statistic of Interest

One of the great advantages of the bootstrap approach is that it can be
applied in almost all situations. No complicated mathematical calculations are required.

Performing a bootstrap analysis in R entails only two steps.

1. We must create a function that computes the statistic of interest.
2. We use the `boot()` function, which is part of the boot library, to perform the bootstrap by repeatedly sampling observations from the data set with replacement.

To illustrate the use of the bootstrap on this data, we must first create
a function, `alpha.fn()`, which takes as input the *(X, Y)* data.

```{r}
alpha.fn=function(data,index){
 X=data$X[index]
 Y=data$Y[index]
 return((var(Y)-cov(X,Y))/(var(X)+var(Y)-2*cov(X,Y)))
}
```

This function returns an estimate for $\alpha$ based on the observations indexed by the argum   ent index. For instance, the following command tells R to estimate $\alpha$ using all 100 observations.

```{r}
library(ISLR)
alpha.fn(Portfolio,1:100)
```

The next command uses the `sample()` function to randomly select 100 observations from the range 1 to 100, with replacement.

This is equivalent to constructing a new bootstrap data set and recomputing $\hat{\alpha}$ based on the new data set.

```{r}
set.seed(1)
alpha.fn(Portfolio, sample(100, 100, replace=T))
```

We can implement a bootstrap analysis by performing this command many times, recording all of the corresponding estimates for $\alpha$, and computing the resulting standard deviation. However, the `boot()` function automates this approach.

Below we produce R=1000 bootstrap estimates for $\alpha$

```{r}
library(boot)
boot(Portfolio, alpha.fn, R=1000)
```

The final output shows that using the original data, $\hat{\alpha}$ = 0.5758, and that the bootstrap estimate for $SE(\hat{\alpha})$ is 0.0886.


### Estimating the Accuracy of a Linear Regression Model

The bootstrap approach can be used to assess the variability of the coefficient estimates and predictions from a statistical learning method.

Here we use the bootstrap approach in order to assess the variability of the estimates for $\beta_0$ and $\beta_1$, the intercept and slope terms for the linear regression model that uses `horsepower` to predict `mpg` in the `Auto` data set.

We will compare the estimates obtained using the bootstrap to those obtained using the formulas for $SE(\hat{\beta_0})$ and $SE(\hat{\beta_1})$

We first create a simple function, `boot.fn()`, which takes in the `Auto` data set as well as a set of indices for the observations, and returns the intercept and slope estimates for the linear regression model.

We then apply this function to the full set of 392 observations in order to compute the estimates of $\hat{\beta_0}$ and $\hat{\beta_1}$ on the entire data set using the usual linear regression coefficient estimate formulas.

```{r}
boot.fn=function(data,index)
 return(coef(lm(mpg~horsepower,data=data,subset=index)))
boot.fn(Auto,1:392)
```

The `boot.fn()` function can also be used in order to create bootstrap estimates for the intercept and slope terms by randomly sampling from among the observations with replacement.

### Examples

```{r}
set.seed(1)
boot.fn(Auto,sample(392,392,replace=T))
boot.fn(Auto,sample(392,392,replace=T))
boot(Auto,boot.fn,1000)
```

This indicates that the bootstrap estimate for $SE(\hat{\beta_0})$ is 0.86, and that the bootstrap estimate for $SE(\hat{\beta_1})$ is 0.0074.

Standard formulas can be used to compute the standard errors for the regression coefficients in a linear model. These can be obtained using the `summary()` function.

```{r}
summary(lm(mpg~horsepower,data=Auto))$coef
```

The standard error estimates for $\hat{\beta_0}$ and $\hat{\beta_1}$ obtained using the formulas from Section 3.1.2 are 0.717 for the intercept and 0.0064 for the slope. Interestingly, these are somewhat different from the estimates obtained using the bootstrap. Does this indicate a problem with the bootstrap? In fact, it suggests the opposite. For example, they depend on the unknown parameter $\sigma^2$, the noise variance.We then estimate $\sigma^2$ using the RSS. Now although the formula for the standard errors do not
rely on the linear model being correct, the estimate for $\sigma^2$, does. 

1. Since there is a non-linear relationship in the data, and so the residuals from a linear fit will be inflated, and so will $\sigma^2$.
2. The standard formulas assume (somewhat unrealistically) that the $x_i$ are fixed, and all the variability comes from the variation in the errors $\epsilon_i$. The bootstrap approach does not rely on any of these assumptions, and so it is likely giving a more accurate estimate of the standard errors of $\hat{\beta_0}$ and $\hat{\beta_1}$.


### Quadratic model of standard linear regression

It is likely giving a more accurate estimate of the standard errors of $\hat{\beta_0}$ and $\hat{\beta_1}$ than is the `summary()` function.

Below we compute the bootstrap standard error estimates and the standard linear regression estimates that result from fitting the quadratic model to the data.

Since this model provides a good fit to the data, there is now a better correspondence between the bootstrap estimates and the standard estimates of $SE(\hat{\beta_0})$, $SE(\hat{\beta_1})$ and $SE(\hat{\beta_2})$.

```{r}
boot.fn=function(data,index)
 coefficients(lm(mpg~horsepower+I(horsepower^2),data=data,subset=index))
set.seed(1)
boot(Auto,boot.fn,1000)
summary(lm(mpg~horsepower+I(horsepower^2),data=Auto))$coef
```

