---
title: "Predict Wage with Splines"
output: html_document
---

## Synopsis

To predict and model wages using Splines. In order to fit regression splines, we will fit it by constructing an appropriate matrix of basis functions.

## Analysis

### Fitting wage to age using a regression spline

```{r}
# Loading the wage packages
library(ISLR)
attach(Wage)

# Load splines package
library(splines)

agelims <- range(age)
age.grid <- seq(from=agelims[1], to=agelims[2])
fit <- lm(wage~bs(age, knots=c(25,40,60)), data=Wage)
pred <- predict(fit, newdata=list(age=age.grid), se=T)

plot(age, wage, col="gray")
lines(age.grid ,pred$fit, lwd=2)
lines(age.grid ,pred$fit+2*pred$se, lty="dashed")
lines(age.grid, pred$fit-2*pred$se, lty="dashed")
```

### Specify knots and percentiles

Here we have prespecified knots at ages 25, 40, and 60. This produces a spline with six basis functions. (Recall that a cubic spline with three knots has seven degrees of freedom; these degrees of freedom are used up by an intercept, plus six basis functions)

We could also use the df option to produce a spline with knots at uniform quantiles of the data.

```{r}
dim(bs(age,knots=c(25,40,60))) # 3 knots
dim(bs(age,df=6)) # 6 basic functions
attr(bs(age,df=6),"knots")
```

In this case R chooses knots at ages 33.8, 42.0, and 51.0, which correspond to the 25th, 50th, and 75th percentiles of `age`.

The `degree` argument allows us to fit splines of any degree, rather than the default degree of 3 (which yields a cubic spline).


### Natural Spline

We fit a natural spline with four degrees of freedom.

```{r}
fit2 <- lm(wage~ns(age, df=4), data=Wage)
pred2 <- predict(fit2, newdata=list(age=age.grid), se=T)

# Add the natural spline in the existing plot
plot(age, wage, col="gray")
lines(age.grid ,pred$fit, lwd=2)
lines(age.grid ,pred$fit+2*pred$se, lty="dashed")
lines(age.grid, pred$fit-2*pred$se, lty="dashed")

lines(age.grid, pred2$fit, col="red", lwd=2) # The natural spline in red
```

### Smoothing Spline

```{r}
plot(age,wage,xlim=agelims,cex=.5,col="darkgrey")
title("Smoothing Spline")

fit <- smooth.spline(age,wage,df=16) # choose 16 degree of freedom
fit2 <- smooth.spline(age,wage,cv=TRUE) # use Cross-validation
fit2$df

lines(fit,col="red",lwd=2)
lines(fit2,col="blue",lwd=2)
legend("topright",legend=c("16 DF","6.8 DF"),col=c("red","blue"),lty=1,lwd=2,cex=.8)
```

Notice that in the first call to smooth.spline(), we specified df=16. The function then determines which value of $\lambda$ leads to 16 degrees of freedom.

In the second call to smooth.spline(), we select the smoothness level by crossvalidation;
this results in a value of $\lambda$ that yields 6.8 degrees of freedom.

### Local Regression

We use the loess() function to pefrom local regression

```{r}
plot(age, wage, xlim=agelims, cex=.5, col="darkgrey")
title("Local Regression")

# Local Regression
fit <- loess(wage~age, span=.2, data=Wage)
fit2 <- loess(wage~age, span=.5, data=Wage)

lines(age.grid, predict(fit, data.frame(age=age.grid)), col="red", lwd=2)
lines(age.grid, predict(fit2, data.frame(age=age.grid)), col="blue", lwd=2)
legend("topright", legend=c("Span=0.2", "Span=0.5"), 
       col=c("red","blue"), lty=1, lwd=2, cex=.8)
```

Here we have performed local linear regression using spans of 0.2 and 0.5: that is, each neighborhood consists of 20% or 50% of the observations. The larger the span, the smoother the fit.


